Analyzing X-ray images in Python with scikit-image

### 2016-08-20 skimage_tomo.tex
Add suggestions for visualization / advanced library sections

The benefits of \texttt{scikit-image} for image processing come not only
from the features of the package alone, but also from the rich
environment surrounding scientific Python \citep{Oliphant2007, Perez2011}.
Fig.~\ref{fig:ecosystem} illustrates how several components of this
ecosystem combine into a sophisticated image processing workflow.

@@

While several interpreters are available to execute Python
instructions and scripts interactively, the most popular in the scientific world is
IPython \citep{Perez2007, Rossant2015}. IPython is an
advanced interpreter, which integrates syntax highlighting, text auto-completion,
a debugger,
introspection and profiling methods, and online help. Several
Integrated Development Environments (IDEs) include IPython,
together with other components such as a text editor. Notable examples
include Spyder (Fig.~\ref{fig:spyder}), PyCharm, and Visual Studio Code.

@@

The Jupyter notebook \citep{Kluyver2016} is a web application that grew
out of the IPython project. Jupyter notebooks provide an interactive development
environment within a web browser, where live code can be enriched
by explanatory text, equations and visualizations (FIGURE).
(jni: FURTHER TEXT ABOUT IMPACT OF JUPYTER NOTEBOOKS)

\paragraph{NumPy arrays} are the cornerstone of the Scientific Python
ecosystem, and of \texttt{scikit-image} operations in particular.
Cropping or downsampling an image, or retrieving pixels corresponding to
a given label in a segmentation are all NumPy ``one-liners''. To
illustrate the compactness of NumPy code, consider modifying pixel values below a
threshold.  This operation can be written as
\begin{lstlisting}
im[im < 0.5] = 0
\end{lstlisting}
exploiting the possibility of indexing arrays with boolean arrays, also called
\emph{masking}. Also, \texttt{NumPy} uses memory sparingly and avoids
making new copies of arrays whenever possible, which is a important
requirement when dealing with the gigabyte-sized images of tomography.
For example, cropping a subvolume as follows does not create a copy of the
original array
\begin{lstlisting}
sub_volume = im[100:-100, 100:-100, 100:-100]
\end{lstlisting}
but instead refers to the correct memory offsets in the original.

@@

Visualizing images is an important component of the image processing
workflow, in order to verify the final result and to adjust the
parameters of intermediate processing operations.
\texttt{matplotlib} \citep{Hunter2007} is the most popular 2D plotting
library of the Python ecosystem. It can be used to visualize 2D data such
as color or grayscale images, and 1D data such as contour lines, outlines
of segmented regions, histograms of gray levels, etc. Although
\texttt{matplotlib} has simple 3D plotting capabilities, we
recommend using the \texttt{mayavi} module \citep{Ramachandran2011}
for applications requiring advanced 3D visualization, such as tomography. 
\texttt{mayavi} is based on the VTK toolkit. It exposes a simple API for
visualizing data passed as \texttt{numpy} arrays. For more advanced
visualizations, a large majority of VTK capabilities can be accessed
thanks to a pipeline API.
\texttt{mayavi} offers a good trade-off between simplicity of use for
common operations, and more advanced possibilities, including responsive
visualizations.

@@

\texttt{scikit-image} is only one Python module that can be used for data
processing, among many others. A very popular module is
\texttt{scikit-learn} \citep{Pedregosa2011}, a Python module for machine
learning using \texttt{NumPy} arrays. If local features of an image
(such as local statistics of gray levels, or geometric points of
interest), or features of segmented objects (e.g. geometrical and
intensity characteristics of segmented particles) are extracted with
functions from \texttt{skimage.feature} (see Fig.~\ref{fig:ecosystem}), it
is then possible to use a \emph{classification} algorithm of
\texttt{scikit-learn}, either for labeling pixels (a segmentation task)
or to classify whole images or objects that have already been segmented.
The interoperability between these packages is ensured by the common use
of \texttt{NumPy} arrays, so that the tasks described above are
implemented with a few lines of code.

>>>>

The benefits of \texttt{scikit-image} for image processing come not only
from the features of the package alone, but also from the rich
environment of Scientific Python \citep{Oliphant2007, Perez2011}.
Fig.~\ref{fig:ecosystem} illustrates how several components of this
ecosystem combine into a sophisticated image processing workflow.

@@

While several interpreters are available to execute interactively Python
instructions and scripts, the most popular one in the scientific world is
the IPython interpreter \citep{Perez2007, Rossant2015}. IPython is an
advanced interpreter, which integrates syntax highlighting, text auto-completion,
a debugger,
introspection and profiling methods, and online help. Several
Integrated Development Environments (IDEs) include IPython,
together with other components such as a text editor. Notable examples
include Spyder (Fig.~\ref{fig:spyder}), PyCharm, and Visual Studio Code.

@@

The Jupyter notebook \citep{Kluyver2016} is a web application that grew
from the IPython project. Jupyter notebooks provide an interactive development
environment within a web browser, where live code can be enriched
by explanatory text, equations and visualizations (FIGURE).
(jni: FURTHER TEXT ABOUT IMPACT OF JUPYTER NOTEBOOKS)

\paragraph{NumPy arrays} are the cornerstone of the Scientific Python
ecosystem, and of \texttt{scikit-image} operations in particular.
Cropping or downsampling an image, or retrieving pixels corresponding to
a given label in a segmentation are all one-liners NumPy code. To
illustrate the compactness of NumPy code, modifying pixels values below a
threshold can be written as
\begin{lstlisting}
im[im < 0.5] = 0
\end{lstlisting}
using the possibility to index arrays with boolean arrays, called
\emph{masking}. Also, \texttt{NumPy} uses memory sparingly and avoids
making new copies of arrays whenever possible, which is a important
requirement when dealing with the gigabyte-sized images of tomography.
For example, cropping a subvolume as follows does not create a copy of the
original array:
\begin{lstlisting}
sub_volume = im[100:-100, 100:-100, 100:-100]
\end{lstlisting}

@@

Visualizing images is an important component of the image processing
workflow, used to inspect the final result and to adjust the
parameters of intermediate processing operations.
\texttt{matplotlib} \citep{Hunter2007} is the most popular 2D plotting
library of the Python ecosystem. It can be used to visualize 2D data such
as color or grayscale images, and 1D data such as contour lines, outlines
of segmented regions, histograms of gray levels, etc. Although
\texttt{matplotlib} has simple 3D plotting capabilities, we
recommend using the \texttt{mayavi} module \citep{Ramachandran2011}
for applications requiring advanced 3D visualization, such as tomography. 
\texttt{mayavi} is based on the VTK toolkit. It exposes a simple API for
visualizing data passed as \texttt{numpy} arrays. For more advanced
visualizations, a large majority of VTK capabilities can be accessed
through mayavi's pipeline API.
\texttt{mayavi} offers a good trade-off between simplicity of use for
common operations, and accessibility to more sophisticated
capabilities such as responsive
visualizations.

@@

\texttt{scikit-image} is only one Python module that can be used for data
processing, among many others. A very popular module is
\texttt{scikit-learn} \citep{Pedregosa2011}, a Python module for machine
learning using \texttt{NumPy} arrays. Local features of an image
(such as local statistics of gray levels, or geometric points of
interest) or features of segmented objects (e.g. geometrical and
intensity characteristics of segmented particles) can be extracted with
functions from \texttt{skimage.feature} (see Fig.~\ref{fig:ecosystem}). It
is then possible to use a \emph{classification} algorithm from
\texttt{scikit-learn} to label pixels (a segmentation task)
or to classify whole images or objects that have already been segmented.
The interoperability between these packages is ensured by the common use
of \texttt{NumPy} arrays, so that the tasks described above are
implemented with a few lines of code.
